---
title: "Exploring the Star Wars Universe"
author: "Katelyn"
output: git_hub
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
library(repurrrsive) #this package contains the sw_films data
library(magrittr) #this package gives us additional pipe options
library(purrr)
library(scales)
trilogies <- factor(
  c("Prequels: Episode I-III", 
    "Originals: Episode IV-VI", 
    "Sequels: Episode VII"), 
  levels = c("Prequels: Episode I-III", 
             "Originals: Episode IV-VI", 
             "Sequels: Episode VII")) #will be used later
```

## Extracting Information from Different Data Sets  

Here is information about the fist 7 Star Wars films:

```{r eval = FALSE}
View(sw_films) 
```

We are going to explore the data contained in several lists similar to this one (and the previously explore `sw_people`), combining skills from all of our previous R code learning experiences.  

How do the following two codes compare?

```{r}
sw_films[[4]][["title"]]
sw_films %>% pluck(4,"title")
```

>they are both ways to grab data, giving you the same answer just two ways possablities.  (good to know both in the case you need to read someones code chunk.)

Suppose we want to pull out just the titles as a character vector, select the correct code (comment out the rest) do perform this action.

```{r}
#comment out the incorrect codes
#sw_films %>% map("title")
sw_films %>% map_chr("title")
#sw_films %>% map_dfc("title")
```

Suppose we want to apply a function to count the number of specific kinds of ships and vehicles in our data.  

Notice that for each film, the "starships" vector contains links to information on those starships (though note this data is out of date and should is linked at swapi.dev, not swapi.co). 
```{r}
sw_films[[1]][["starships"]]
```

So if we can count the number of webpage links that would tell us the number of starships that appear in that movie.  Here are three different ways to count the number of urls under `starships`. Can you think of another? (it is ok if you can't). Compare and contrast how the three codes work differently to do the same thing.

```{r}
sw_films %>% map(., "starships") %>% map_dbl(~length(.))
map_dbl(sw_films, ~length(.x$starships))
sw_films %>% map_dbl(., ~length(.x$starships))
```

> the first one grabs the data and returns a list or vector then grom it grabbs the length of the numbers in starships. where as the next one starts with looking in the data to grab numbers for the x. data set under starships where as the last one withh look into the sw_films data and pulled outt he length with in the straships section numerically. 



## Evaluating Hyperdrive in the Star Wars Episodes

We will use the third method from the previous section to extract out the information we want from `sw_films`.  For each row, specify if we should use a regular `map()`, `map_dbl()`, or `map_chr()`.

**NOTE** Sometimes code like this gets a little finicky in R if you try to run it with `Ctrl + Enter`.  Instead, use the code chunk green arrow to run the whole code chunk or highlight all of the code and then use the shortcut to run it.  

```{r}
sw_ships <- sw_films %>% {
  tibble(
    title = map_chr(., "title"), #character
    episode = map_dbl(., "episode_id"), #numeric
    starships = map_dbl(., ~length(.x$starships)), #numeric
    vehicles = map_dbl(., ~length(.x$vehicles)), #numeric
    planets = map_dbl(., ~length(.x$planets)) #numeric
  )}
sw_ships
```

Let's do a bit more data cleaning to 1) assign the Trilogy classification to each episode, 2) calculate the total number of starships (which have hyperdrive) and vehicles (which do not have hyperdrive), and 3) calculate the proportion of total ships that have hyperdrive. Fill in the missing codes.

```{r}
#use the pipe that will assign this information back to sw_ships
sw_ships %<>%  
  #create a new variable called trilogy
  mutate(trilogy = case_when(episode %in% 1:3 ~ trilogies[1],
                             episode %in% 4:6 ~ trilogies[2],
                             episode %in% 7 ~ trilogies[3])) %>% 
  #create a new variable called total_ships which adds vehicles and starships together
  mutate(total_ships = vehicles + starships) %>%  
  #create a new variable called prop that calculate the percent hyperdrive
  mutate(prop = starships / total_ships * 100) 
```


Now, let's make a plot examining how often hyperdrive ships appear in each episode.  You can see the image `bar_hyp.png` to see what the graph should look like and fill in the missing codes:

```{r}
sw_ships %>% 
  #be sure to order titles by order/episode
  ggplot(aes(y = order_by(title, desc(episode)), x=prop)) + 
  #we want bars but our data is already summarized!
  geom_col(aes(fill = trilogy)) + 
  labs(
    title = "The Rise of Hyperdrive",
    subtitle = "Percentage of Ships with Hyperdrive Capability"
  ) +
  #you may need to install `scales` package if you haven't already
  scale_x_continuous(labels = scales::label_percent(scale=1)) +
  theme_minimal() +
  #what aesthetic do we modify to change the bar color
  scale_fill_viridis_d(end = 0.8) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank())

```

We can also look at a plot to see if there is a correlation between the total number of ships and the number with hyperdrive (starships). You can see the image `scatter_hyp.png` to see what the graph should look like and fill in the missing codes:

```{r}
sw_ships %>% 
  ggplot(aes(x = total_ships, y = starships)) +
  #make points
  geom_point(aes(color = trilogy)) +
  #fit a smooth model
  geom_smooth(method = "lm") +
  #what does geom_text() do?
  geom_text(aes(label = title), vjust = -1, 
            hjust = "inward", size = 2.75) +
  labs(title = "Hyperdrive Correlations",
       subtitle = "The Number of Ships with Hyperdrive vs Total Ships") +
  theme_minimal() +
  #what aesthetic do we want to modify the color of points?
  scale_colour_viridis_d(end = 0.8) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank()) 
```

What do you notice about the use of hyperdrive type vehicles in the episodes?  

> the original and he Prequils use about the same amount of hyperdrive ships as one another t least for the most part. but the sequels dont match up as well.


## The Physical Features of Star Wars Characters

Recall the data for "people" in Star Wars:

```{r eval = FALSE}
View(sw_people)
```


We want to extract out `name`, `height`, and `mass` as `character` vectors (for now, we have to deal with some issues in height and weight later to change them into double type vectors) and keep `films` as a list for now.  Fill in the correct `map` type functions for each one.

```{r}
sw_peeps <- sw_people %>%  {
  tibble(
    name = map_chr(., "name"),  #character
    height = map_chr(., "height"), #character
    mass = map_chr(., "mass"), #character
    films = map(., "films") #list
  )}
sw_peeps
```

Notice that the `films` column contains lists of urls for each film reference.  Let's pull out that same information from the `sw_films` data to have the `title` of the episode and the `url` as a `character` vector, and the episode number as a numeric value.  Fill in the correct `map` type functions.  

```{r}
film_names <- sw_films %>% {
  tibble(
    episode_id = map_dbl(., "episode_id"), #double
    episode_name = map_chr(., "title"), #character
    url = map_chr(., "url") #character
  )}
film_names
```


Now we can finish cleaning up our data by doing the following: 1) turn `height` and `mass` into numeric vectors; 2) match the `films`/`urls` to their `episode_names` and assign that using the correct pipe function back to `sw_peeps`.

```{r}
sw_peeps %>%  #be careful running this code multiple times since it is using the assignment pip - you may want to change to the %>% pipe first and to make sure you have the right output and then change back to the assignment pipe
  #use a function from readr to extract the numbers and replace "unknown" with na
  mutate(height = parse_number(height, na = "unknown"),
         mass = parse_number(mass, na = "unknown")) %>%
  #unnest the lists in films
  unnest(cols = c("films")) %>% 
  #join the film data with episodes names to the people data
  left_join(film_names, by = c("films" = "url")) %>% 
  #remove the `films` url from the data frame
  select(-films) %>% 
  #add the variable trilogy
  mutate(trilogy = case_when(episode_id %in% 1:3 ~ trilogies[1],
                             episode_id %in% 4:6 ~ trilogies[2],
                             episode_id %in% 7 ~ trilogies[3]))
sw_peeps
```


We can now create a plot of height and mass by trilogy group, you can see the final product in `scater_people.png`")`:

```{r}
sw_peeps %>% 
  filter(name != "Jabba Desilijic Tiure") %>% #major outlier removed
  #map the correct aesthetics
  ggplot(aes(x = height, y = mass, color = "Trilogy Group"))+
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Height of Character (cm)",
       y = "Mass of Character (kg)",
       color = "Trilogy Group",
       title = "Character Characteristics in Star Wars") +
  theme_minimal() +
  scale_color_viridis_d(end = 0.8) 
```



## OPTIONAL CHALLENGE PROBLEM
Your professor wants to use `purrr` to try and generate a height and mass scatterplot for each episode, but I don't want to type out all that code.  Here is where I got so far, but I am not convinced this is the most sophisticated or effective way to do this.  Do some research and see if you can find a way to put this process into production!

```{r}
plots_sw <- sw_peeps %>% 
  nest(data = !episode_name) %>% 
  mutate(plot = map2(data, episode_name, 
                     ~ggplot(data = .x, aes(y = mass, 
                                            x = height)) + 
                      geom_point() + 
                      geom_smooth(method = "lm", se = FALSE) + 
                      ggtitle(.y)))
```


```{r}
print(plots_sw$plot)
```

